---
title: "ReproducibleResearch_PA1"
author: "Michael Nechay"
date: "Saturday, January 17, 2015"
output: html_document
---

Load data and preprocess date column

```{r}
rawData <- read.csv("activity.csv")
rawData$date <- as.Date(rawData$date, "%Y-%m-%d")
totalSteps <- aggregate(steps ~ date, data = rawData, sum, na.rm = TRUE)
head(rawData)
```

###What is the mean total number of steps taken per day?

histogram of total number of steps taken each day:

```{r}
hist(totalSteps$steps, main="Total number of steps taken", ylab="Frequency", xlab="Number of Steps", col="red")
```
###Mean and median number of steps taken each day.

```{r}
mean(totalSteps$steps, na.rm = T)
median(totalSteps$steps, na.rm = T)
```

###Time series plot of the average number of steps taken (averaged across all days)

```{r}
intervalSteps <- aggregate(steps ~ interval, data = rawData, mean, na.rm = TRUE)
plot(intervalSteps$interval, intervalSteps$steps, type="l", ylab ="mean steps", xlab = "interval")
```

total number of missing data
```{r}
sum(as.numeric(is.na(rawData$steps)))
```

###Filling in missing data - Strategy:

Replace each NA with the mean for the corresponding 5-minute interval. 

```{r}
newData <- rawData
for (i in 1:nrow(rawData) )
{
  if ( is.na(rawData[i,1]) ) {
    interval <- rawData[i,3]
    correspondingRow <- interval / 5 + 1
    newData[i,1] <- intervalSteps[correspondingRow,2]
  }
}
```

###Histogram of the total number of steps after data is filled in:

```{r}
newTotalSteps <- aggregate(steps ~ date, data = newData, sum, na.rm = TRUE)
hist(newTotalSteps$steps, main="Total number of steps taken", ylab="Frequency", xlab="Number of Steps", col="red")

mean(newTotalSteps$steps, na.rm = T)
median(newTotalSteps$steps, na.rm = T)
```

###Are there differences in activity patterns between weekdays and weekends?

```{r}
newData$posixdate <- as.POSIXlt(newData$date)

#weekends
weekends <- newData[which(newData$posixdate$wday == 0 | newData$posixdate$wday == 6),]
intervalStepsWeekends <- aggregate(steps ~ interval, data = weekends, mean, na.rm = TRUE)

#weekdays
weekdays <- newData[which(newData$posixdate$wday > 0 & newData$posixdate$wday < 6),]
intervalStepsWeekdays <- aggregate(steps ~ interval, data = weekdays, mean, na.rm = TRUE)
```

histogram of average number of steps taken per 5-minute interval, colored by weekend/weekday:
```{r}
par(mfrow=c(1,1))
plot(intervalStepsWeekdays$interval, intervalStepsWeekdays$steps, type="l", ylab ="mean steps", xlab = "interval")
lines(intervalStepsWeekends$interval, intervalStepsWeekends$steps, col="blue")
legend("topright", legend=c("Weekdays", "Weekends"), col=c("black", "blue"), lty=1, lwd=2)
```
